---
title: "LaTeX Table Tests"
author: "Benjamin Nutter"
date: "October 1, 2015"
output: pdf_document
header-includes:
 - \usepackage{longtable}
 - \usepackage[dvipsnames,table]{xcolor}
 - \usepackage{array}
 - \usepackage{arydshln}
 - \usepackage{amssymb}
 - \usepackage{graphicx}
 - \makeatletter
 - \newcommand*\vdashline{\rotatebox[origin=c]{90}{$\dabar@\dabar@\dabar@$}}
 - \makeatother
---

```{r}
library(dplyr)
library(pixiedust)

mtcars2 <- mtcars[1:10, ]

Hmisc::label(mtcars2$mpg) <- "Gas Mileage"
Hmisc::label(mtcars2$cyl) <- "Cylinders"
Hmisc::label(mtcars2$disp) <- "Displacement"
Hmisc::label(mtcars2$hp) <- "Horse Power"
Hmisc::label(mtcars2$drat) <- "Rear Axle Ratio"
Hmisc::label(mtcars2$wt) <- "Weight"
Hmisc::label(mtcars2$qsec) <- "1/4 mile time"
Hmisc::label(mtcars2$vs) <- "V/S"
Hmisc::label(mtcars2$am) <- "Transmission"
Hmisc::label(mtcars2$gear) <- "Forward Gears"
Hmisc::label(mtcars2$carb) <- "Carburetors"

custom_head <- rbind(names(mtcars2), Hmisc::label(mtcars2)) %>%
  as.data.frame(stringsAsFactors = FALSE)

custom_foot <- rbind(vapply(mtcars2, mean, numeric(1)),
                     vapply(mtcars2, sd, numeric(1))) %>%
  as.data.frame(stringsAsFactors = FALSE)
```

## Longtable

```{r}
custom_interfoot <- data.frame("To Be Continued", 
                               "", "", "", "", "", "",
                               "", "", "", "")

(x <- dust(mtcars2) %>%
  redust(custom_head, part = "head") %>%
  redust(custom_foot, part = "foot") %>%
  redust(custom_interfoot, part = "interfoot") %>%
  sprinkle_table(round = 2, longtable = 4) %>%
  sprinkle(bg = "gray", part = "head") %>%
  sprinkle(bg = "lightgray", part = "foot") %>%
  sprinkle(bg = "lightgray", part = "interfoot") %>%
  sprinkle_print_method("latex"))
```

\begin{center}
\begin{longtable}{|c|c|c|c|}
\caption{A simple longtable example}\\

\textbf{First entry} & \textbf{Second entry} & \textbf{Third entry} & \textbf{Fourth entry} \\
\endfirsthead

\textbf{First entry} & \textbf{Second entry} & \textbf{Third entry} & \textbf{Fourth entry} \\
\endhead

\multicolumn{4}{r}{Continued on next page} \\
\endfoot

\endlastfoot
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\ 1 & 2 & 3 & 4 \\
\end{longtable}
\end{center}
\end{document}