---
title: "Sprinkles"
author: "Benjamin Nutter"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: no
    number_section: yes
    toc: yes
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Dustbunnies}
  \usepackage[utf8]{inputenc}
---

Sprinkles are used to modify the appearance of a table. The table below shows the currently available and planned sprinkles (in light green and yellow, respectively) as well as the implementations for the printing methods.

Black indicates that a sprinkle will have no impact on the output for that method.

Red indicates that a sprinkle is not yet functional for that method.

Green indicates that a sprinkle is currently functional for that method.


```{r, echo=FALSE}
Sprinkles <- read.csv(system.file("sprinkles.csv", package = "pixiedust"),
                        stringsAsFactors=FALSE)
Sprinkles[,-1] <- lapply(Sprinkles[-1], trimws)
```

```{r, echo=FALSE}
library(pixiedust)
red <- "#A50026"
lightgreen <- "#A6DBA0"
green <- "#006837"
yellow <- "#FFFFBF"

x <- dust(Sprinkles[, -2]) + 
  sprinkle(row = which(Sprinkles$implemented == ""), 
           col = "dustbunny", bg = yellow) + 
  sprinkle(row = which(Sprinkles$implemented == "x"),
           col = "dustbunny", bg = lightgreen)
x$body$bg <- ifelse(!is.na(x$body$bg),
                    x$body$bg,
                    ifelse(x$body$value == "",
                           "black",
                           ifelse(x$body$value == "o",
                                  red,
                                  green)))
x$body$value <- ifelse(x$body$value %in% c("x", "o"), "", x$body$value)
  
x + sprinkle_print_method(print_method = "html")

```
